---
title: "Channel_4_dataset"
author: "V_warrier"
date: "02/04/2018"
output:
  html_document: default
  latex_engine: xelatex
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(plyr)
library(reshape2)
data = read.csv("data- originalchannel4 bigdata.csv", header = T, fill = T) #758916 obs
cat("the total number of IDs in this file is", nrow(data))
```

\newpage

# Section 1: Data cleaning and generating variables

\newpage

## data cleanup and recoding variables

```{r}
#remove_repeats
data2 = subset(data, repeat. == 0) #695166

cat("the number of non-repeated participants is", nrow(data2))

#keep males of females
data2 = subset(data2, sex == 1 | sex == 2)#672279

cat("\n")

cat("Of this, the number of participants who are either male or female is",  nrow(data2))

##keep age bound

data2 = subset(data2, age > 15 & age < 90)

cat("\n")

cat("Of this, the number participants after all QC is ",  nrow(data2))

```

\newpage

## Recoding variables
```{r}
#recode AQ
data2[,c(51,57,58,60)] <- lapply(data2[,c(51,57,58,60)], function(x) 
  recode(x,"1" = 1, "2" = 1, "3" = 0, "4" =0 ))

data2[,c(52,53,54,55,56,59)] <- lapply(data2[,c(52,53,54,55,56,59)], function(x) 
   recode(x,"1" = 0, "2" = 0, "3" = 1, "4" =1))

data2$AQ_full = data2$AQ_1 + data2$AQ_2 + data2$AQ_3 + data2$AQ_4 + data2$AQ_5 + data2$AQ_6 + data2$AQ_7 + data2$AQ_8 + data2$AQ_9 + data2$AQ_10

data2$AQ_Z = scale(data2$AQ_full, center = TRUE, scale = TRUE)


#Recode EQ
data2[,c(31, 32, 34, 36, 39)] <- lapply(data2[,c(31, 32, 34, 36, 39)], function(x) 
  recode(x,"1" = 2, "2" = 1, "3" = 0, "4" =0 ))

data2[,c(33, 35, 37, 38, 40 )] <- lapply(data2[,c(33, 35, 37, 38, 40)], function(x) 
  recode(x,"1" = 0, "2" = 0, "3" = 1, "4" =2))

data2$EQ_full = data2$EQ_1 + data2$EQ_2 + data2$EQ_3 + data2$EQ_4 + data2$EQ_5 + data2$EQ_6 + data2$EQ_7 + data2$EQ_8 + data2$EQ_9 + data2$EQ_10

data2$EQ_Z = scale(data2$EQ_full, center = TRUE, scale = TRUE)


#Recode SQ
data2[,c(41,43, 44, 46, 47, 49, 50)] <- lapply(data2[,c(41,43, 44, 46, 47, 49, 50)], function(x) 
  recode(x,"1" = 2, "2" = 1, "3" = 0, "4" =0 ))

data2[,c(42, 45, 48 )] <- lapply(data2[,c(42, 45, 48)], function(x) 
  recode(x,"1" = 0, "2" = 0, "3" = 1, "4" =2))

data2$SQ_full = data2$SQR_1 + data2$SQR_2 + data2$SQR_3 + data2$SQR_4 + data2$SQR_5 + data2$SQR_6 + data2$SQR_7 + data2$SQR_8 + data2$SQR_9 + data2$SQR_10

data2$SQ_Z = scale(data2$SQ_full, center = TRUE, scale = TRUE)


#Recode SPQ

data2[,c(21:30)] <- lapply(data2[,c(21:30)], function(x) 
  recode(x,"1" = 3, "2" = 2, "3" = 1, "4" =0 ))

data2$SPQ_full = data2$SPQ_1 + data2$SPQ_2 + data2$SPQ_3 + data2$SPQ_4 + data2$SPQ_5 + data2$SPQ_6 + data2$SPQ_7 + data2$SPQ_8 + data2$SPQ_9 + data2$SPQ_10


data2$SPQ_Z = scale(data2$SPQ_full, center = TRUE, scale = TRUE)

data2 = data2[!is.na(data2$AQ_full),]

```


\newpage

## Defining cases

```{R}
#Define cases

#define cases based on all different options
data2$autism = ifelse(data2$diagnosis_0 == "2" | data2$diagnosis_1 == "2"| data2$diagnosis_3 == "2" | data2$diagnosis_4 == "2" | data2$diagnosis_5 == "2" | data2$diagnosis_6 == "2" | data2$diagnosis_7 == "2" | data2$diagnosis_8 == "2" | data2$autism_diagnosis_0 > 0 | data2$autism_diagnosis_1 > 0 | data2$autism_diagnosis_2 > 0, 1, 0)

data2$autism[is.na(data2$autism)] <- 0


#define cases based only on the autism criteria
data2$autism2 = ifelse(data2$autism_diagnosis_0 > 0 | data2$autism_diagnosis_1 > 0 | data2$autism_diagnosis_2 > 0, 1, 0)

data2$autism2[is.na(data2$autism2)] <- 0

#count the number of cases


a =  nrow(subset(data2, autism == 1))

cat("the number of autism cases is (broad criteria)", a)

cat("\n")

b = nrow(subset(data2, autism2 == 1))

cat("the number of autism cases is (narrow criteria)", b)


#define schizophrenia
data2$schizophrenia = ifelse(data2$diagnosis_0 == "7" | data2$diagnosis_1 == "7"| data2$diagnosis_3 == "7" | data2$diagnosis_4 == "7" | data2$diagnosis_5 == "7" | data2$diagnosis_6 == "7" | data2$diagnosis_7 == "7" | data2$diagnosis_8 == "7" | data2$autism_diagnosis_0 == "0" | data2$autism_diagnosis_1 == "0" | data2$autism_diagnosis_2 == "0", 1, 0)

data2$schizophrenia[is.na(data2$schizophrenia)] <- 0

#define ocd
data2$ocd = ifelse(data2$diagnosis_0 == "6" | data2$diagnosis_1 == "6"| data2$diagnosis_3 == "6" | data2$diagnosis_4 == "6" | data2$diagnosis_5 == "6" | data2$diagnosis_6 == "6" | data2$diagnosis_7 == "6" | data2$diagnosis_8 == "6" | data2$autism_diagnosis_0 == "0" | data2$autism_diagnosis_1 == "0" | data2$autism_diagnosis_2 == "0", 1, 0)

data2$ocd[is.na(data2$ocd)] <- 0


#define ld
data2$ld = ifelse(data2$diagnosis_0 == "5" | data2$diagnosis_1 == "5"| data2$diagnosis_3 == "5" | data2$diagnosis_4 == "5" | data2$diagnosis_5 == "5" | data2$diagnosis_6 == "5" | data2$diagnosis_7 == "5" | data2$diagnosis_8 == "5" | data2$autism_diagnosis_0 == "0" | data2$autism_diagnosis_1 == "0" | data2$autism_diagnosis_2 == "0", 1, 0)

data2$ld[is.na(data2$ld)] <- 0

#define depression
data2$depression = ifelse(data2$diagnosis_0 == "4" | data2$diagnosis_1 == "4"| data2$diagnosis_3 == "4" | data2$diagnosis_4 == "4" | data2$diagnosis_5 == "4" | data2$diagnosis_6 == "4" | data2$diagnosis_7 == "4" | data2$diagnosis_8 == "4" | data2$autism_diagnosis_0 == "0" | data2$autism_diagnosis_1 == "0" | data2$autism_diagnosis_2 == "0", 1, 0)

data2$depression[is.na(data2$depression)] <- 0


#define bipolar
data2$bipolar = ifelse(data2$diagnosis_0 == "3" | data2$diagnosis_1 == "3"| data2$diagnosis_3 == "3" | data2$diagnosis_4 == "3" | data2$diagnosis_5 == "3" | data2$diagnosis_6 == "3" | data2$diagnosis_7 == "3" | data2$diagnosis_8 == "3" | data2$autism_diagnosis_0 == "0" | data2$autism_diagnosis_1 == "0" | data2$autism_diagnosis_2 == "0", 1, 0)

data2$bipolar[is.na(data2$bipolar)] <- 0

#define adhd
data2$adhd = ifelse(data2$diagnosis_0 == "1" | data2$diagnosis_1 == "1"| data2$diagnosis_3 == "1" | data2$diagnosis_4 == "1" | data2$diagnosis_5 == "1" | data2$diagnosis_6 == "1" | data2$diagnosis_7 == "1" | data2$diagnosis_8 == "6" | data2$autism_diagnosis_0 == "0" | data2$autism_diagnosis_1 == "0" | data2$autism_diagnosis_2 == "0", 1, 0)

data2$ocd[is.na(data2$ocd)] <- 0


```

\newpage

## Defining other variables

```{r}
#Define other variables

data2$STEM = ifelse(data2$occupation  == "3" | data2$occupation == "5" | data2$occupation == "21" , 1, 0)
data2$STEM[is.na(data2$STEM)] <- 0
data2$STEM = ifelse(data2$occupation == "26", 0, data2$STEM)

STEM = subset(data2, STEM == "1")

cat("No of participants in STEM occupation is", nrow(STEM))

cat("\n")


###update country region
data2$countryregion = ifelse(data2$countryregion == "14", 0, data2$countryregion)

###update education

data2$education = ifelse(data2$education == "5", 0, data2$education)

###update handedness

data2$handedness = ifelse(data2$handedness == "4", 0, data2$handedness)


```


\newpage

## Generating D-scores and brain types

```{R}
#This is Wheelwright's method: https://www.ncbi.nlm.nih.gov/pubmed/16473340

controls = subset(data2, autism == "0")
cases = subset(data2, autism == "1")

meanSQ = mean(controls$SQ_full)
meanEQ = mean(controls$EQ_full)

data2$SQ_full_standardized_w = (data2$SQ_full - meanSQ)/20
data2$EQ_full_standardized_w = (data2$EQ_full - meanEQ)/20
data2$wheelwrightD = data2$SQ_full_standardized_w - data2$EQ_full_standardized_w

data2$dpercentile = ntile(data2$wheelwrightD, 100)

data2$braintype = ifelse(data2$dpercentile < 2.5, "1",
                 ifelse(between(data2$dpercentile, 2.499, 35), "2",
                    ifelse(between(data2$dpercentile, 34.99, 65), "3", 
                           ifelse(between(data2$dpercentile, 64.99 , 97.5), "4", "5"))))

controls = subset(data2, autism == "0") #redefining cases and controls again so downstream analysis can include braintype
cases = subset(data2, autism == "1")

```


\newpage

# seperating by sex

```{R}

males_controls = subset(controls, sex == "1")
females_controls = subset(controls, sex == "2")

dim(males_controls)

dim(females_controls)


males_cases = subset(cases, sex == "1")
females_cases = subset(cases, sex == "2")

dim(males_cases)
dim(females_cases)
```

\newpage


################################################################################################
################################################################################################
################################################################################################






# Section 2: Sex, age and geographical location differences

\newpage

## Sex differences in controls

```{R}
#Basic statistics

# Sex differences
t.test(controls$AQ_full ~ controls$sex)
ggplot(controls, aes(x=AQ_full, colour=as.character(sex))) +   geom_density() + theme_minimal()



t.test(controls$EQ_full ~ controls$sex)
ggplot(controls, aes(x=EQ_full, colour=as.character(sex))) +   geom_density() + theme_minimal()


t.test(controls$SQ_full ~ controls$sex)
ggplot(controls, aes(x=SQ_full, colour=as.character(sex))) +   geom_density() + theme_minimal()



t.test(controls$SPQ_full ~ controls$sex)
ggplot(controls, aes(x=SPQ_full, colour=as.character(sex))) +   geom_density() + theme_minimal()


controls2 = controls[,c ("sex", "AQ_full", "EQ_full", "SQ_full", "SPQ_full" )]
controls2 = na.omit(controls2)
controls2_melt =  melt(controls2, id.vars=c("sex"))
ddply(controls2_melt, c("sex", "variable"), summarise,
      mean = mean(value), sd = sd(value),
      sem = sd(value)/sqrt(length(value)))


```

\newpage

## Sex differences in cases

```{R}

t.test(cases$AQ_full ~ cases$sex)
ggplot(cases, aes(x=AQ_full, colour=as.character(sex))) +   geom_density() + theme_minimal()


t.test(cases$EQ_full ~ cases$sex)
ggplot(cases, aes(x=EQ_full, colour=as.character(sex))) +   geom_density() + theme_minimal()


t.test(cases$SQ_full ~ cases$sex)
ggplot(cases, aes(x=SQ_full, colour=as.character(sex))) +   geom_density() + theme_minimal()



t.test(cases$SPQ_full ~ cases$sex)
ggplot(cases, aes(x=SPQ_full, colour=as.character(sex))) +   geom_density() + theme_minimal()


cases2 = cases[,c ("sex", "AQ_full", "EQ_full", "SQ_full", "SPQ_full" )]
cases2 = na.omit(cases2)
cases2_melt =  melt(cases2, id.vars=c("sex"))
ddply(cases2_melt, c("sex", "variable"), summarise,
      mean = mean(value), sd = sd(value),
      sem = sd(value)/sqrt(length(value)))

```

\newpage

## Quantifying differences in effect sizes

```{R}

summary(lm(AQ_full ~ as.character(sex) + as.character(autism) + as.character(sex):as.character(autism), data = data2))
summary(lm(EQ_full ~ as.character(sex) + as.character(autism) + as.character(sex):as.character(autism), data = data2))
summary(lm(SQ_full ~ as.character(sex) + as.character(autism) + as.character(sex):as.character(autism), data = data2))
summary(lm(SPQ_full ~ as.character(sex) + as.character(autism) + as.character(sex):as.character(autism), data = data2))



```


\newpage

## Correlations between AQ, EQ, SQ, and SPQ in the controls data
```{R}

cor.test(controls$AQ_full, controls$EQ_full)

cor.test(controls$AQ_full, controls$SQ_full)

cor.test(controls$AQ_full, controls$SPQ_full)

cor.test(controls$EQ_full, controls$SQ_full)

cor.test(controls$EQ_full, controls$SPQ_full)

cor.test(controls$SQ_full, controls$SPQ_full)

```



\newpage

## Age differences in controls

```{R}
#Age differences
cor.test(controls$AQ_full, controls$age)
ggplot(controls, aes(AQ_full, age)) + geom_smooth(method = "lm") + theme_minimal()

cor.test(controls$EQ_full, controls$age)
ggplot(controls, aes(EQ_full, age)) + geom_smooth(method = "lm") + theme_minimal()

cor.test(controls$SQ_full, controls$age)
ggplot(controls, aes(SQ_full, age)) + geom_smooth(method = "lm") + theme_minimal()

cor.test(controls$SPQ_full, controls$age)
ggplot(controls, aes(SPQ_full, age)) + geom_smooth(method = "lm") + theme_minimal()

```



\newpage

## Age differences in cases

```{R}
#Age differences
cor.test(cases$AQ_full, cases$age)
ggplot(cases, aes(AQ_full, age)) + geom_smooth(method = "lm") + theme_minimal()

cor.test(cases$EQ_full, cases$age)
ggplot(cases, aes(EQ_full, age)) + geom_smooth(method = "lm") + theme_minimal()

cor.test(cases$SQ_full, cases$age)
ggplot(cases, aes(SQ_full, age)) + geom_smooth(method = "lm") + theme_minimal()

cor.test(cases$SPQ_full, cases$age)
ggplot(cases, aes(SPQ_full, age)) + geom_smooth(method = "lm") + theme_minimal()

```


\newpage

## Correlation with education, controls

```{R}
#Correlation with education

cor.test(controls$AQ_full, controls$education)
ggplot(controls, aes(education, AQ_full)) + geom_smooth(method = "lm") + theme_minimal()

cor.test(controls$EQ_full, controls$education)
ggplot(controls, aes(education, EQ_full)) + geom_smooth(method = "lm") + theme_minimal()

cor.test(controls$SQ_full, controls$education)
ggplot(controls, aes(education, SQ_full)) + geom_smooth(method = "lm") + theme_minimal()

cor.test(controls$SPQ_full, controls$education)
ggplot(controls, aes(education, SPQ_full)) + geom_smooth(method = "lm") + theme_minimal()
```



\newpage

## Correlation with education, cases

```{R}
#Correlation with education

cor.test(cases$AQ_full, cases$education)
ggplot(cases, aes(education, AQ_full)) + geom_smooth(method = "lm") + theme_minimal()

cor.test(cases$EQ_full, cases$education)
ggplot(cases, aes(education, EQ_full)) + geom_smooth(method = "lm") + theme_minimal()

cor.test(cases$SQ_full, cases$education)
ggplot(cases, aes(education, SQ_full)) + geom_smooth(method = "lm") + theme_minimal()

cor.test(cases$SPQ_full, cases$education)
ggplot(cases, aes(education, SPQ_full)) + geom_smooth(method = "lm") + theme_minimal()
```


\newpage

## Geographical differences in the four measures, controls

```{R}
m = aov(data=controls, AQ_full~as.character(countryregion))
anova(m)

m = aov(data=controls, EQ_full~as.character(countryregion))
anova(m)

m = aov(data=controls, SQ_full~as.character(countryregion))
anova(m)

m = aov(data=controls, SPQ_full~as.character(countryregion))
anova(m)


controls2 = controls[,c ("sex", "countryregion", "AQ_full", "EQ_full", "SQ_full", "SPQ_full" )]

controls2_melt =  melt(controls2, id.vars=c("sex", "countryregion"))

head(controls2_melt)

alpha = ddply(controls2_melt, c("sex", "countryregion", "variable"), summarise,
      mean = mean(value), sd = sd(value),
      sem = sd(value)/sqrt(length(value)))

alpha

```

\newpage


################################################################################################
################################################################################################
################################################################################################







# Section 3: STEM differences

\newpage

## STEM differences

```{R}
#STEM differences
t.test(controls$AQ_full ~ controls$STEM)
ggplot(controls, aes(x=AQ_full, colour=as.character(STEM))) +   geom_density() + theme_minimal()

t.test(controls$EQ_full ~ controls$STEM)
ggplot(controls, aes(x=EQ_full, colour=as.character(STEM))) +   geom_density() + theme_minimal()

t.test(controls$SQ_full ~ controls$STEM)
ggplot(controls, aes(x=SQ_full, colour=as.character(STEM))) +   geom_density() + theme_minimal()

t.test(controls$SPQ_full ~ controls$STEM)
ggplot(controls, aes(x=SPQ_full, colour=as.character(STEM))) +   geom_density() + theme_minimal()


controls2 = controls[,c ("sex", "STEM", "AQ_full", "EQ_full", "SQ_full", "SPQ_full" )]
controls2 =  controls2[!is.na(controls2$AQ_full),]

controls2_melt =  melt(controls2, id.vars=c("sex", "STEM"))

head(controls2_melt)

alpha = ddply(controls2_melt, c("sex", "STEM", "variable"), summarise,
      mean = mean(value), sd = sd(value),
      sem = sd(value)/sqrt(length(value)))

alpha
```

\newpage

## STEM regression analyses
**AQ analysis**

```{R}
summary(lm(AQ_full ~ STEM + as.character(countryregion) + as.character(handedness) + as.character(sex) + education + age, data = controls))


summary(lm(EQ_full ~ STEM + as.character(countryregion) + as.character(handedness) + as.character(sex) + education + age, data = data2))


summary(lm(SQ_full ~ STEM + as.character(countryregion) + as.character(handedness) + as.character(sex) + education + age, data = controls))



summary(lm(SPQ_full ~ STEM + as.character(countryregion) + as.character(handedness) + as.character(sex) + education + age, data = controls))

```


\newpage


################################################################################################
################################################################################################
################################################################################################



# Section 4: Predicting AQ from other variables in controls



\newpage

```{R}

summary(lm(AQ_full ~ as.character(handedness) + as.character(sex) + as.character(countryregion) + education + age + as.character(STEM), data = controls ))

summary(lm(AQ_full ~ as.character(handedness) + as.character(sex) + as.character(countryregion) + education + age + as.character(STEM) + wheelwrightD, data = controls ))

summary(lm(AQ_full ~ as.character(handedness) + as.character(sex) + as.character(countryregion) + education + age + as.character(STEM) + wheelwrightD + SPQ_full, data = controls ))

summary(lm(AQ_full ~ as.character(handedness) + as.character(sex) + as.character(countryregion) + education + age + as.character(STEM) + SQ_full + EQ_full + SPQ_full, data = controls ))

```



\newpage


################################################################################################
################################################################################################
################################################################################################



# Section 4: Case-control analyses



\newpage


## Case-control STEM
```{R}
cases_STEM = nrow(subset(cases, STEM == "1"))

cases_nonSTEM = nrow(subset(cases, STEM == "0"))

controls_STEM = nrow(subset(controls, STEM == "1"))

controls_nonSTEM = nrow(subset(controls, STEM == "0"))

STEMdiff <- matrix(c(cases_STEM,controls_STEM,cases_nonSTEM,controls_nonSTEM), ncol=2)
colnames(STEMdiff) <- c('STEM', 'nonSTEM')
row.names(STEMdiff) <- c('Cases', 'Controls')

STEMdiff

chisq.test(STEMdiff) 

summary(glm(STEM ~ as.character(autism) + as.character(sex) + education + age, data = data2))

```



\newpage



## Case-control AQ, EQ, SQ, and SPQ
```{R}

males = subset(data2, sex == "1")
females = subset(data2, sex == "2")
t.test(males$AQ_full ~ males$autism)
ggplot(males, aes(x=AQ_full, colour=as.character(autism))) +   geom_density() + theme_minimal()

t.test(males$EQ_full ~ males$autism)
ggplot(males, aes(x=EQ_full, colour=as.character(autism))) +   geom_density() + theme_minimal()

t.test(males$SQ_full ~ males$autism)
ggplot(males, aes(x=SQ_full, colour=as.character(autism))) +   geom_density() + theme_minimal()

t.test(males$SPQ_full ~ males$autism)
ggplot(males, aes(x=SPQ_full, colour=as.character(autism))) +   geom_density() + theme_minimal()


t.test(females$AQ_full ~ females$autism)
ggplot(females, aes(x=AQ_full, colour=as.character(autism))) +   geom_density() + theme_minimal()

t.test(females$EQ_full ~ females$autism)
ggplot(females, aes(x=EQ_full, colour=as.character(autism))) +   geom_density() + theme_minimal()

t.test(females$SQ_full ~ females$autism)
ggplot(females, aes(x=SQ_full, colour=as.character(autism))) +   geom_density() + theme_minimal()

t.test(females$SPQ_full ~ females$autism)
ggplot(females, aes(x=SPQ_full, colour=as.character(autism))) +   geom_density() + theme_minimal()

data2melt = data2[,c ("sex", "autism", "AQ_full", "EQ_full", "SQ_full", "SPQ_full" )]
data2melt =  data2melt[!is.na(data2melt$AQ_full),]
data2melt =  melt(data2melt, id.vars=c("sex", "autism"))


head(data2melt)

ddply(data2melt, c("sex", "autism", "variable"), summarise,
      mean = mean(value), sd = sd(value),
      sem = sd(value)/sqrt(length(value)))


```



\newpage

## Case-control regression analyses**

```{R}

summary(glm(as.numeric(as.character(autism)) ~ as.character(sex) + education + 
    AQ_full + AQ_full:education + AQ_full:as.character(sex) + as.character(sex):education + 
    as.character(countryregion) + as.character(handedness) + 
    age + age:education + AQ_full:age + age:as.character(sex) + STEM, data = data2))




```


\newpage

## Brain type analysis
### Females control percentage

```{R}

#Sex distribution

controls = subset(data2, autism == "0")
fem_controls = subset(controls, sex == "2")

cat("total number of females is", nrow(fem_controls))

extremeE = subset(fem_controls, braintype == "1")
E = subset(fem_controls, braintype == "2")
B = subset(fem_controls, braintype == "3")
S = subset(fem_controls, braintype == "4")
extremeS = subset(fem_controls, braintype == "5")

cat("percentage of females with extreme E is", (nrow(extremeE)/nrow(fem_controls))*100)

cat("percentage of females with E is", (nrow(E)/nrow(fem_controls))*100 )

cat("percentage of females with B is", (nrow(B)/nrow(fem_controls))*100)

cat("percentage of females with S is", (nrow(S)/nrow(fem_controls))*100)

cat("percentage of females with Extreme S is", (nrow(extremeS)/nrow(fem_controls))*100)

```


\newpage

## Brain type analysis
### Males percentage

```{R}

#Sex distribution

controls = subset(data2, autism == "0")
male_controls = subset(controls, sex == "1")

cat("total number of males is", nrow(male_controls))

extremeE = subset(male_controls, braintype == "1")
E = subset(male_controls, braintype == "2")
B = subset(male_controls, braintype == "3")
S = subset(male_controls, braintype == "4")
extremeS = subset(male_controls, braintype == "5")

cat("percentage of males with extreme E is", (nrow(extremeE)/nrow(male_controls))*100)

cat("percentage of males with E is", (nrow(E)/nrow(male_controls))*100 )

cat("percentage of males with B is", (nrow(B)/nrow(male_controls))*100)

cat("percentage of males with S is", (nrow(S)/nrow(male_controls))*100)

cat("percentage of males with Extreme S is", (nrow(extremeS)/nrow(male_controls))*100)

```


\newpage

## Brain type analysis
### Females cases percentage

```{R}

#Sex distribution

cases = subset(data2, autism == "1")
fem_cases = subset(cases, sex == "2")

cat("total number of females is", nrow(fem_cases))

extremeE = subset(fem_cases, braintype == "1")
E = subset(fem_cases, braintype == "2")
B = subset(fem_cases, braintype == "3")
S = subset(fem_cases, braintype == "4")
extremeS = subset(fem_cases, braintype == "5")

cat("percentage of females with extreme E is", (nrow(extremeE)/nrow(fem_cases))*100)

cat("percentage of females with E is", (nrow(E)/nrow(fem_cases))*100 )

cat("percentage of females with B is", (nrow(B)/nrow(fem_cases))*100)

cat("percentage of females with S is", (nrow(S)/nrow(fem_cases))*100)

cat("percentage of females with Extreme S is", (nrow(extremeS)/nrow(fem_cases))*100)

```

\newpage

## Brain type analysis
### Males cases percentage

```{R}

#Sex distribution

cases = subset(data2, autism == "1")
male_cases = subset(cases, sex == "1")

cat("total number of males is", nrow(male_cases))

extremeE = subset(male_cases, braintype == "1")
E = subset(male_cases, braintype == "2")
B = subset(male_cases, braintype == "3")
S = subset(male_cases, braintype == "4")
extremeS = subset(male_cases, braintype == "5")

cat("percentage of females with extreme E is", (nrow(extremeE)/nrow(male_cases))*100)

cat("percentage of females with E is", (nrow(E)/nrow(male_cases))*100 )

cat("percentage of females with B is", (nrow(B)/nrow(male_cases))*100)

cat("percentage of females with S is", (nrow(S)/nrow(male_cases))*100)

cat("percentage of females with Extreme S is", (nrow(extremeS)/nrow(male_cases))*100)

```

\newpage


##Anova with brain types
```{R}

m = aov(data=controls, AQ_full~as.character(braintype))
anova(m)
TukeyHSD((m), "as.character(braintype)")


m = aov(data=controls, SPQ_full~as.character(braintype))
anova(m)
TukeyHSD((m), "as.character(braintype)")

controls2 = controls[,c ("braintype", "AQ_full", "SPQ_full" )]
controls2 =  controls2[!is.na(controls2$AQ_full),]
controls2_melt =  melt(controls2, id.vars=c("braintype"))


ddply(controls2_melt, c("braintype", "variable"), summarise,
      mean = mean(value), sd = sd(value),
      sem = sd(value)/sqrt(length(value)))


```




## Sex differences
```{R}
sexdiff = matrix(c(18230, 241604, 18515, 393930), ncol = 2, byrow = TRUE)

colnames(sexdiff) = c("Cases", "Controls")

rownames(sexdiff) = c("Males", "Females")

sexdiff

chisq.test(sexdiff)

```

\newpage

## Additional analyses
### Other sex

```{R}
data3 = subset(data, repeat. == 0) #695166

data3$autism = ifelse(data3$diagnosis_0 == "2" | data3$diagnosis_1 == "2"| data3$diagnosis_3 == "2" | data3$diagnosis_4 == "2" | data3$diagnosis_5 == "2" | data3$diagnosis_6 == "2" | data3$diagnosis_7 == "2" | data3$diagnosis_8 == "2" | data3$autism_diagnosis_0 > 0 | data3$autism_diagnosis_1 > 0 | data3$autism_diagnosis_2 > 0, 1, 0)

data3$autism[is.na(data3$autism)] <- 0

cases_2 = subset(data3, autism == "1")
controls_2 = subset(data3, autism == "0")

controls1 = subset(controls_2, sex == "1")
controls2 = subset(controls_2, sex == "2")
controls3 = subset(controls_2, sex == "3")

a = nrow(controls1)

b = nrow(controls2)

c = nrow(controls3)

cases1 = subset(cases_2, sex == "1")
cases2 = subset(cases_2, sex == "2")
cases3 = subset(cases_2, sex == "3")

d = nrow(cases1)

e = nrow(cases2)

f = nrow(cases3)

sex_3_way = matrix(c(a, d, b, e, c, f ), ncol = 2, byrow = TRUE)

colnames(sex_3_way) = c("Controls", "Cases")

rownames(sex_3_way) = c("Males", "Females", "Other")

sex_3_way

chisq.test(sex_3_way)

sex_2_way = matrix(c(d+e, a+b, f, c ), ncol = 2, byrow = TRUE)

colnames(sex_2_way) = c("Cases", "Controls")

rownames(sex_2_way) = c("Binary", "Non-binary")

sex_2_way

chisq.test(sex_2_way)

```


\newpage

## ANOVA with handedness in controls

```{R}
#ANOVA with handedness
###AQ###

cat("handedness AQ results")
m = aov(data=controls, AQ_full~as.character(handedness))
anova(m)
TukeyHSD((m), "as.character(handedness)")

tky = as.data.frame(TukeyHSD(m)$"as.character(handedness)")
tky$pair = rownames(tky)
tky$`p adj` = ifelse(tky$`p adj` == 0, 2.2e-16, tky$`p adj`)


# Plot pairwise TukeyHSD comparisons and color by significance level
ggplot(tky, aes(colour=cut(`p adj`, c(0, 0.01, 0.05, 1), 
                           label=c("p<0.01","p<0.05","Non-Sig")))) +
  geom_hline(yintercept=0, lty="11", colour="grey30") +
  geom_errorbar(aes(pair, ymin=lwr, ymax=upr), width=0.2) +
  geom_point(aes(pair, diff)) +
  labs(colour="") +labs(x = "Difference", y = "Comparisons") + theme(legend.position="none") + theme_minimal()


###EQ###

cat("handedness EQ results")
m = aov(data=controls, EQ_full~as.character(handedness))
anova(m)
TukeyHSD((m), "as.character(handedness)")

tky = as.data.frame(TukeyHSD(m)$"as.character(handedness)")
tky$pair = rownames(tky)
tky$`p adj` = ifelse(tky$`p adj` == 0, 2.2e-16, tky$`p adj`)


# Plot pairwise TukeyHSD comparisons and color by significance level
ggplot(tky, aes(colour=cut(`p adj`, c(0, 0.01, 0.05, 1), 
                           label=c("p<0.01","p<0.05","Non-Sig")))) +
  geom_hline(yintercept=0, lty="11", colour="grey30") +
  geom_errorbar(aes(pair, ymin=lwr, ymax=upr), width=0.2) +
  geom_point(aes(pair, diff)) +
  labs(colour="") +labs(x = "Difference", y = "Comparisons") + theme(legend.position="none") + theme_minimal()


###SQ###
cat("handedness SQ results")
m = aov(data=controls, SQ_full~as.character(handedness))
anova(m)
TukeyHSD((m), "as.character(handedness)")

tky = as.data.frame(TukeyHSD(m)$"as.character(handedness)")
tky$pair = rownames(tky)
tky$`p adj` = ifelse(tky$`p adj` == 0, 2.2e-16, tky$`p adj`)


# Plot pairwise TukeyHSD comparisons and color by significance level
ggplot(tky, aes(colour=cut(`p adj`, c(0, 0.01, 0.05, 1), 
                           label=c("p<0.01","p<0.05","Non-Sig")))) +
  geom_hline(yintercept=0, lty="11", colour="grey30") +
  geom_errorbar(aes(pair, ymin=lwr, ymax=upr), width=0.2) +
  geom_point(aes(pair, diff)) +
  labs(colour="") +labs(x = "Difference", y = "Comparisons") + theme(legend.position="none") + theme_minimal()

###SPQ##
cat("handedness SPQ results")
m = aov(data=controls, SPQ_full~as.character(handedness))
anova(m)
TukeyHSD((m), "as.character(handedness)")

tky = as.data.frame(TukeyHSD(m)$"as.character(handedness)")
tky$pair = rownames(tky)
tky$`p adj` = ifelse(tky$`p adj` == 0, 2.2e-16, tky$`p adj`)


# Plot pairwise TukeyHSD comparisons and color by significance level
ggplot(tky, aes(colour=cut(`p adj`, c(0, 0.01, 0.05, 1), 
                           label=c("p<0.01","p<0.05","Non-Sig")))) +
  geom_hline(yintercept=0, lty="11", colour="grey30") +
  geom_errorbar(aes(pair, ymin=lwr, ymax=upr), width=0.2) +
  geom_point(aes(pair, diff)) +
  labs(colour="") +labs(x = "Difference", y = "Comparisons") + theme(legend.position="none") + theme_minimal()

```


\newpage

# Generating plots

```{R}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

data2$category = ifelse(data2$sex == 1 & data2$autism == 0, "control_males", "NA")
data2$category = ifelse(data2$sex == 2 & data2$autism == 0, "control_females", data2$category)
data2$category = ifelse(data2$sex == 2 & data2$autism == 1, "autism_females", data2$category)
data2$category = ifelse(data2$sex == 1 & data2$autism == 1, "autism_males", data2$category)

#CDFfigure
cdf_plot = ggplot(data2, aes(wheelwrightD, colour = category)) + stat_ecdf() +xlab("Dscore") + ylab("Cumulative frequency") + theme_classic()

#Case-control by sex Figure

a = ggplot(data2, aes(x=AQ_full, colour=as.character(category))) +   geom_density(adjust = 4) + theme_classic() + xlab("AQ-10 scores") + scale_x_continuous(breaks=seq(0,11,2))
b = ggplot(data2, aes(x=EQ_full, colour=as.character(category))) +   geom_density(adjust = 2) + theme_classic() + xlab("EQ-10 scores") + scale_x_continuous(breaks=seq(0,41,2))
c = ggplot(data2, aes(x=SQ_full, colour=as.character(category))) +   geom_density(adjust = 2) + theme_classic() + xlab("SQ-10 scores") + scale_x_continuous(breaks=seq(0,41,2))
d = ggplot(data2, aes(x=SPQ_full, colour=as.character(category))) +   geom_density(adjust = 2) + theme_classic() + xlab("SPQ-10 scores") + scale_x_continuous(breaks=seq(0,31,5))
case_control_plot = multiplot(a, b, c, d, cols=2)


#STEM figure
controls2 = controls[!is.na(controls$STEM),]
controls2$category = ifelse(controls2$sex == 1 & controls2$STEM == 0, "nonSTEM_males", "NA")
controls2$category = ifelse(controls2$sex == 2 & controls2$STEM == 0, "nonSTEM_females", controls2$category)
controls2$category = ifelse(controls2$sex == 1 & controls2$STEM == 1, "STEM_males", controls2$category)
controls2$category = ifelse(controls2$sex == 2 & controls2$STEM == 1, "STEM_females", controls2$category)

a = ggplot(controls2, aes(x=AQ_full, colour=as.character(category))) +   geom_density(adjust = 4) + theme_classic() + xlab("AQ-10 scores") + scale_x_continuous(breaks=seq(0,11,2))
b = ggplot(controls2, aes(x=EQ_full, colour=as.character(category))) +   geom_density(adjust = 2) + theme_classic() + xlab("EQ-10 scores") + scale_x_continuous(breaks=seq(0,41,2))
c = ggplot(controls2, aes(x=SQ_full, colour=as.character(category))) +   geom_density(adjust = 2) + theme_classic() + xlab("SQ-10 scores") + scale_x_continuous(breaks=seq(0,41,2))
d = ggplot(controls2, aes(x=SPQ_full, colour=as.character(category))) +   geom_density(adjust = 2) + theme_classic() + xlab("SPQ-10 scores") + scale_x_continuous(breaks=seq(0,31,5))
stem_diff_plot = multiplot(a, b, c, d, cols=2)


```






