---
title: "Channel_4_validation"
author: "Varun Warrier"
date: "11 May 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



This is a file to validate 7 out of the 10 findings:

(1) Females on average will score higher on empathy 
(2) That males on average will score higher on systemizing
(3) That there will be a trade-off between E and S. 
(4) That there are 5 distinct 'brain types' based on the difference (or D-score) between E and S. These brain types are Type B (balanced, where E=S), Type E (where E>S), Type S (where S>E), Extreme Type E (E>>S) and Extreme Type S (S>>E). 
(5) That more females will have a brain of Type E, and 
(6) that more males will have a brain of Type S.
(7) That more people with autism will have a brain of Extreme Type S.

\newpage

#Step 1: read dataset and create variables
```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(plyr)
library(reshape2)

data1 = fread("C:/Users/Varun/Desktop/Manuscripts 2018/Channel_4/replication_mu.txt")
data1$sex = ifelse(data1$sex == "1", "Male", "Female")
data1$autism  = ifelse(data1$autism == "1", "Autism", "Control")
data1$category = ifelse(data1$autism == "Autism" & data1$sex == "Male", "Autism_male", NA)
data1$category = ifelse(data1$autism == "Autism" & data1$sex == "Female", "Autism_female", data1$category)
data1$category = ifelse(data1$autism == "Control" & data1$sex == "Female", "Control_female", data1$category)
data1$category = ifelse(data1$autism == "Control" & data1$sex == "Male", "Control_male", data1$category)

data1= data1[!is.na(data1$age),]

data1 = subset(data1, age > 17 & age < 89)

a =  nrow(subset(data1, category == "Control_male"))

cat("the number of control males is", a)

a =  nrow(subset(data1, category == "Control_female"))

cat("the number of control females is", a)

a =  nrow(subset(data1, category == "Autism_male"))

cat("the number of Autism males is", a)

a =  nrow(subset(data1, category == "Autism_female"))

cat("the number of Autism females is", a)
```

\newpage

#Generate D scores
```{r}
controls = subset(data1, autism == "Control")
cases = subset(data1, autism == "Autism")

meanSQ = mean(controls$SQ_total)
meanEQ = mean(controls$EQ_total)

data1$SQ_full_standardized_w = (data1$SQ_total - meanSQ)/50
data1$EQ_full_standardized_w = (data1$EQ_total - meanEQ)/80
data1$wheelwrightD = data1$SQ_full_standardized_w - data1$EQ_full_standardized_w

data1$dpercentile = ntile(data1$wheelwrightD, 100)

data1$braintype = ifelse(data1$dpercentile < 2.5, "1",
                 ifelse(between(data1$dpercentile, 2.499, 35), "2",
                    ifelse(between(data1$dpercentile, 34.99, 65), "3", 
                           ifelse(between(data1$dpercentile, 64.99 , 97.5), "4", "5"))))



```

\newpage

#Run analysis and generate figures: predictions 1 and 2
```{r}
# Sex differences
t.test(controls$EQ_total ~ controls$sex)
t.test(controls$SQ_total ~ controls$sex)


t.test(cases$EQ_total ~ cases$sex)
t.test(cases$SQ_total ~ cases$sex)

cor.test(scale(controls$EQ_total), scale(controls$SQ_total))

data2 = data1[,c ("category", "EQ_total", "SQ_total" )]
data2 = na.omit(data2)
data2_melt =  melt(data2, id.vars=c("category"))
ddply(data2_melt, c("category", "variable"), summarise,
      mean = mean(value), sd = sd(value),
      sem = sd(value)/sqrt(length(value)))


```

\newpage

## Brain type analysis
### Females control percentage

```{R}

#Sex distribution

fem_controls = subset(data1, category == "Control_female")

cat("total number of females is", nrow(fem_controls))

extremeE = subset(fem_controls, braintype == "1")
E = subset(fem_controls, braintype == "2")
B = subset(fem_controls, braintype == "3")
S = subset(fem_controls, braintype == "4")
extremeS = subset(fem_controls, braintype == "5")

cat("percentage of females with extreme E is", (nrow(extremeE)/nrow(fem_controls))*100)

cat("percentage of females with E is", (nrow(E)/nrow(fem_controls))*100 )

cat("percentage of females with B is", (nrow(B)/nrow(fem_controls))*100)

cat("percentage of females with S is", (nrow(S)/nrow(fem_controls))*100)

cat("percentage of females with Extreme S is", (nrow(extremeS)/nrow(fem_controls))*100)

```


\newpage

## Brain type analysis
### Males percentage

```{R}

#Sex distribution


male_controls = subset(data1, category == "Control_male")

cat("total number of males is", nrow(male_controls))

extremeE = subset(male_controls, braintype == "1")
E = subset(male_controls, braintype == "2")
B = subset(male_controls, braintype == "3")
S = subset(male_controls, braintype == "4")
extremeS = subset(male_controls, braintype == "5")

cat("percentage of males with extreme E is", (nrow(extremeE)/nrow(male_controls))*100)

cat("percentage of males with E is", (nrow(E)/nrow(male_controls))*100 )

cat("percentage of males with B is", (nrow(B)/nrow(male_controls))*100)

cat("percentage of males with S is", (nrow(S)/nrow(male_controls))*100)

cat("percentage of males with Extreme S is", (nrow(extremeS)/nrow(male_controls))*100)

```


\newpage

## Brain type analysis
### Females cases percentage

```{R}

#Sex distribution

fem_cases = subset(data1, category == "Autism_female")

cat("total number of females is", nrow(fem_cases))

extremeE = subset(fem_cases, braintype == "1")
E = subset(fem_cases, braintype == "2")
B = subset(fem_cases, braintype == "3")
S = subset(fem_cases, braintype == "4")
extremeS = subset(fem_cases, braintype == "5")

cat("percentage of females with extreme E is", (nrow(extremeE)/nrow(fem_cases))*100)

cat("percentage of females with E is", (nrow(E)/nrow(fem_cases))*100 )

cat("percentage of females with B is", (nrow(B)/nrow(fem_cases))*100)

cat("percentage of females with S is", (nrow(S)/nrow(fem_cases))*100)

cat("percentage of females with Extreme S is", (nrow(extremeS)/nrow(fem_cases))*100)

```

\newpage

## Brain type analysis
### Males cases percentage

```{R}

#Sex distribution

male_cases = subset(data1, category == "Autism_male")

cat("total number of males is", nrow(male_cases))

extremeE = subset(male_cases, braintype == "1")
E = subset(male_cases, braintype == "2")
B = subset(male_cases, braintype == "3")
S = subset(male_cases, braintype == "4")
extremeS = subset(male_cases, braintype == "5")

cat("percentage of males with extreme E is", (nrow(extremeE)/nrow(male_cases))*100)

cat("percentage of males with E is", (nrow(E)/nrow(male_cases))*100 )

cat("percentage of males with B is", (nrow(B)/nrow(male_cases))*100)

cat("percentage of males with S is", (nrow(S)/nrow(male_cases))*100)

cat("percentage of males with Extreme S is", (nrow(extremeS)/nrow(male_cases))*100)

```




